Modelling low Mach number stellar hydrodynamics with MAESTROeX

### 2019-10-02 astronum_2019_MAESTROeX/astronum-maestroex.tex
some updates

\maestroex~is currently able to model spherically symmetric systems, however it has no support for rotation. We wish to model convection in the interiors of massive stars, and it is known that these stars often have non-negligible rotational frequencies and that this rotation can have have significant effect on mixing at convective shell boundaries. We are therefore currently exploring several possible ways of implementing rotation in \maestroex.

@@

\MarginPar{I feel like someone who knows what they're talking about should write/seriously edit this section}
The latest generation of supercomputers to come online (e.g.~Summit, soon Perlmutter and Frontier) rely heavily on GPUs in order to achieve high performance while keeping energy consumption to a minimum (GPUs consume considerably less power per flop than CPUs). This trend is set to continue: supercomputers are moving to new architectures in order to increase performance rather than simply adding more and more CPUs. In order for HPC codes to be able to exploit these latest machines, it is therefore becoming increasingly necessary to port codes to run on GPUs.

We have begun porting \maestroex~to GPUs, leveraging the GPU capabilities built into \amrex~and following the lessons learned porting \castro~to GPUs. In the \amrex~framework, routines can be ported to GPUs by simply inserting a few macros into the code. The boilerplate code to required to compile this for GPU is then generated by a custom preprocessor prior to compilation. CUDA managed memory is used to take care of data transfers between the CPU and GPU, so no explicit copies of data to/from the CPU and GPU are required, significantly reducing the extra code required to port the code. One of the key emphases of the \amrex~approach to GPUs is to maintain performance portability: porting code to GPUs should not come at the expense of the code's performance on CPU-only machines. Because GPU-specific code is created by the preprocessor prior to compilation, it is not necessary to maintain separate versions of the code for different architectures. By changing the compiler flags, the same bit of code can be compiled for serial, MPI-only, OpenMP-only, MPI+OpenMP, GPU-only or MPI+GPU.   
In order to achieve the best performance on the GPU, we require some additional optimization of the computational kernals. However, we find that these optimizations often improve the performance when the code is run on the CPU only as well. 
More details to appear in a forthcoming \amrexastro~GPU paper.

>>>>

\maestroex~is currently able to model spherically symmetric systems, however it has no support for rotation. \MarginPar{I would say it has no support for rotation that breaks the spherical symmetry} We wish to model convection in the interiors of massive stars, and it is known that these stars often have non-negligible rotational frequencies and that this rotation can have have significant effect on mixing at convective shell boundaries. We are therefore currently exploring several possible ways of implementing rotation in \maestroex.

@@

\MarginPar{I feel like someone who knows what they're talking about should write/seriously edit this section}
The latest generation of supercomputers to come online (e.g.~ OLCF Summit, soon NERSC Perlmutter and OLCF Frontier) rely heavily on GPUs in order to achieve high performance while keeping energy consumption to a minimum (GPUs consume considerably less power per flop than CPUs). This trend is set to continue: supercomputers are moving to new architectures in order to increase performance rather than simply adding more and more CPUs. In order for HPC codes to be able to exploit these latest machines, it is therefore becoming increasingly necessary to port codes to run on GPUs.

We have begun porting \maestroex~to GPUs, leveraging the GPU capabilities built into \amrex~and following the lessons learned porting \castro~to GPUs. In the \amrex~framework, routines can be ported to GPUs by simply inserting a few macros into the code. The boilerplate code to required to compile this for GPU is then generated by a custom preprocessor prior to compilation. CUDA managed memory is used to take care of data transfers between the CPU and GPU, so no explicit copies of data to/from the CPU and GPU are required, significantly reducing the extra code required to port the code. One of the key emphases of the \amrex~approach to GPUs is to maintain performance portability: porting code to GPUs should not come at the expense of the code's performance on CPU-only machines. Because GPU-specific code is created by the preprocessor prior to compilation, it is not necessary to maintain separate versions of the code for different architectures. By changing the compiler flags, the same bit of code can be compiled for serial, MPI-only, OpenMP-only, MPI+OpenMP, GPU-only or MPI+GPU.   
In order to achieve the best performance on the GPU, we require some additional optimization of the computational kernels. However, we find that these optimizations often improve the performance when the code is run on the CPU only as well. 
More details to appear in a forthcoming \amrexastro~GPU paper.
